---
title: "Tobacco Susceptibility Measurement"
author: "Scott I. Donaldson, PhD"
date: "08/24/2026"
editor_options: 
  chunk_output_type: console
---
```{r Packages}
#Packages
library(readxl)
library(dplyr)
library(psych)
library(mirt)
library(ggplot2)
library(mokken)
library(WrightMap)
library(lavaan)
library(semTools)
library(tidyverse)
```

```{r Import IPSOS data}
IPSOS_Youth <- read_excel("Youth Ipsos 2026_Data for Scott.xlsx")

```
```{r Import PATH data}
# Load PATH Wave 8 Youth data
load("Wave8_Youth_DATA/DS8002/36498-8002-Data.rda")
youth <- da36498.8002

# Load PATH Wave 8 Adult / Young Adult data
load("Wave8_YA_DATA/DS8001/36498-8001-Data.rda")
adult <- da36498.8001

# Create young adult sample from adult data
ya <- adult[adult$R08R_A_AGECAT6 == "(1) 1 = 18 to 24 years old", ]

```

#Recode PATH Variables
```{r Recode PATH susceptibility variables}
# Recode susceptibility: 1 = lowest susceptibility, 4 = highest susceptibility
recode_sus <- function(x) {
  x_num <- as.numeric(x)
  ifelse(is.na(x_num), NA, 5 - x_num)
}

# Harmonized PATH Youth susceptibility variables
PATH_youth_sus <- youth %>%
  dplyr::transmute(
    source = "PATH",
    dataset = "PATH youth",
    group = "Youth",

    # Nicotine pouches
    np_curious  = recode_sus(R08_YA1103NP),
    np_nextyear = recode_sus(R08_YA1206NP),
    np_soon     = recode_sus(R08_YA1104NP),
    np_friend   = recode_sus(R08_YA1105NP),

    # Cigarettes
    cig_curious  = recode_sus(R08_YC1103),
    cig_nextyear = recode_sus(R08_YC1206),
    cig_soon     = recode_sus(R08_YC1104),
    cig_friend   = recode_sus(R08_YC1105),

    # E-cigarettes / ENDS
    ecig_curious  = recode_sus(R08_YV1103),
    ecig_nextyear = recode_sus(R08_YV1206),
    ecig_soon     = recode_sus(R08_YV1104),
    ecig_friend   = recode_sus(R08_YV1105),

    # Hookah
    hookah_curious  = recode_sus(R08_YH1103),
    hookah_nextyear = recode_sus(R08_YH1206),
    hookah_soon     = recode_sus(R08_YH1104),
    hookah_friend   = recode_sus(R08_YH1105),

    # Snus
    snus_curious  = recode_sus(R08_YU1103),
    snus_nextyear = recode_sus(R08_YU1206),
    snus_soon     = recode_sus(R08_YU1104),
    snus_friend   = recode_sus(R08_YU1105),

    # Smokeless tobacco
    smkls_curious  = recode_sus(R08_YS1103),
    smkls_nextyear = recode_sus(R08_YS1206),
    smkls_soon     = recode_sus(R08_YS1104),
    smkls_friend   = recode_sus(R08_YS1105),

    # Traditional cigars
    cigar_curious  = recode_sus(R08_YG9010),
    cigar_nextyear = recode_sus(R08_YG9020),
    cigar_soon     = recode_sus(R08_YG9013),
    cigar_friend   = recode_sus(R08_YG9016),

    # Cigarillos
    cigarillo_curious  = recode_sus(R08_YG9011),
    cigarillo_nextyear = recode_sus(R08_YG9021),
    cigarillo_soon     = recode_sus(R08_YG9014),
    cigarillo_friend   = recode_sus(R08_YG9017),

    # Filtered cigars
    fcigar_curious  = recode_sus(R08_YG9012),
    fcigar_nextyear = recode_sus(R08_YG9022),
    fcigar_soon     = recode_sus(R08_YG9015),
    fcigar_friend   = recode_sus(R08_YG9018)
  )

# Harmonized PATH Young Adult susceptibility variables
PATH_ya_sus <- ya %>%
  dplyr::transmute(
    source = "PATH",
    dataset = "PATH young adult",
    group = "Young adult",

    # Nicotine pouches
    np_curious  = recode_sus(R08_AA1103NP),
    np_nextyear = recode_sus(R08_AA1206NP),
    np_soon     = recode_sus(R08_AA1104NP),
    np_friend   = recode_sus(R08_AA1105NP),

    # Cigarettes
    cig_curious  = recode_sus(R08_AC1103),
    cig_nextyear = recode_sus(R08_AC1206),
    cig_soon     = recode_sus(R08_AC1104),
    cig_friend   = recode_sus(R08_AC1205),

    # E-cigarettes / ENDS
    ecig_curious  = recode_sus(R08_AV1103),
    ecig_nextyear = recode_sus(R08_AV1206),
    ecig_soon     = recode_sus(R08_AV1104),
    ecig_friend   = recode_sus(R08_AV1105),

    # Hookah
    hookah_curious  = recode_sus(R08_AH1103),
    hookah_nextyear = recode_sus(R08_AH1206),
    hookah_soon     = recode_sus(R08_AH1104),
    hookah_friend   = recode_sus(R08_AH1205),

    # Snus
    snus_curious  = recode_sus(R08_AU1103),
    snus_nextyear = recode_sus(R08_AU1206),
    snus_soon     = recode_sus(R08_AU1104),
    snus_friend   = recode_sus(R08_AU1205),

    # Smokeless tobacco
    smkls_curious  = recode_sus(R08_AS1103),
    smkls_nextyear = recode_sus(R08_AS1206),
    smkls_soon     = recode_sus(R08_AS1104),
    smkls_friend   = recode_sus(R08_AS1205),

    # Traditional cigars
    cigar_curious  = recode_sus(R08_AG9010),
    cigar_nextyear = recode_sus(R08_AG9020),
    cigar_soon     = recode_sus(R08_AG9013),
    cigar_friend   = recode_sus(R08_AG9016),

    # Cigarillos
    cigarillo_curious  = recode_sus(R08_AG9011),
    cigarillo_nextyear = recode_sus(R08_AG9021),
    cigarillo_soon     = recode_sus(R08_AG9014),
    cigarillo_friend   = recode_sus(R08_AG9017),

    # Filtered cigars
    fcigar_curious  = recode_sus(R08_AG9012),
    fcigar_nextyear = recode_sus(R08_AG9022),
    fcigar_soon     = recode_sus(R08_AG9015),
    fcigar_friend   = recode_sus(R08_AG9018)
  )

# Combine PATH youth and young adult susceptibility data
PATH_sus_wide <- dplyr::bind_rows(PATH_youth_sus, PATH_ya_sus)

# Pouch-only analytic PATH dataset for earlier analyses
np_all_analytic <- PATH_sus_wide %>%
  dplyr::select(source, dataset, group, np_curious, np_nextyear, np_soon, np_friend) %>%
  dplyr::filter(rowSums(!is.na(dplyr::select(., np_curious, np_nextyear, np_soon, np_friend))) > 0)

# Check analytic sample size by group
table(np_all_analytic$group)

# Check distributions
summary(np_all_analytic)
```
```{r PATH broad susceptibility data}
# Product item sets
product_items <- list(
  np = c("np_curious", "np_nextyear", "np_soon", "np_friend"),
  cig = c("cig_curious", "cig_nextyear", "cig_soon", "cig_friend"),
  ecig = c("ecig_curious", "ecig_nextyear", "ecig_soon", "ecig_friend"),
  hookah = c("hookah_curious", "hookah_nextyear", "hookah_soon", "hookah_friend"),
  snus = c("snus_curious", "snus_nextyear", "snus_soon", "snus_friend"),
  smkls = c("smkls_curious", "smkls_nextyear", "smkls_soon", "smkls_friend"),
  cigar = c("cigar_curious", "cigar_nextyear", "cigar_soon", "cigar_friend"),
  cigarillo = c("cigarillo_curious", "cigarillo_nextyear", "cigarillo_soon", "cigarillo_friend"),
  fcigar = c("fcigar_curious", "fcigar_nextyear", "fcigar_soon", "fcigar_friend")
)

# All broad susceptibility item names
broad_sus_items <- unname(unlist(product_items))

# Check that all items exist in PATH_sus_wide
setdiff(broad_sus_items, names(PATH_sus_wide))

# Create PATH-only broad susceptibility analytic dataset
PATH_broad_sus <- PATH_sus_wide %>%
  dplyr::select(
    source,
    dataset,
    group,
    dplyr::all_of(broad_sus_items)
  )

# Keep respondents with at least one susceptibility item answered
PATH_broad_sus_analytic <- PATH_broad_sus %>%
  dplyr::filter(
    rowSums(!is.na(dplyr::select(., dplyr::all_of(broad_sus_items)))) > 0
  )

# Check sample
table(PATH_broad_sus_analytic$dataset)
summary(PATH_broad_sus_analytic[, broad_sus_items])
```
```{r Broad item descriptive ranges}
# Reproducible item-level descriptives for Results reporting.
# psych::describe() reports excess kurtosis (normal distribution = 0).
broad_item_descriptives <- psych::describe(
  PATH_broad_sus_analytic[, broad_sus_items],
  fast = FALSE
) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("item") %>%
  dplyr::select(item, n, mean, sd, skew, kurtosis)

broad_descriptive_ranges <- broad_item_descriptives %>%
  dplyr::summarise(
    mean_min = min(mean, na.rm = TRUE),
    mean_max = max(mean, na.rm = TRUE),
    sd_min = min(sd, na.rm = TRUE),
    sd_max = max(sd, na.rm = TRUE),
    skew_min = min(skew, na.rm = TRUE),
    skew_max = max(skew, na.rm = TRUE),
    kurtosis_min = min(kurtosis, na.rm = TRUE),
    kurtosis_max = max(kurtosis, na.rm = TRUE)
  ) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), ~ round(.x, 2)))

broad_item_descriptives
broad_descriptive_ranges
```

#Recode IPSOS Youth
```{r Recode}
# Preserve raw IPSOS data
IPSOS_Youth_raw <- IPSOS_Youth

# Identify respondent ID column across IPSOS file versions
ipsos_id_col <- dplyr::case_when(
  "ResponseId" %in% names(IPSOS_Youth_raw) ~ "ResponseId",
  "caseid2" %in% names(IPSOS_Youth_raw) ~ "caseid2",
  TRUE ~ NA_character_
)

if (is.na(ipsos_id_col)) stop("No IPSOS respondent ID column found")

# Sus recode function: 1 = lowest susceptibility, 4 = highest susceptibility
recode_ipsos_sus <- function(x) {
  x_clean <- stringr::str_to_lower(stringr::str_trim(as.character(x)))
  
  dplyr::case_when(
    x_clean %in% c(".", "", "na", "nan") ~ NA_real_,
    stringr::str_detect(x_clean, "definit") & stringr::str_detect(x_clean, "not") ~ 1,
    stringr::str_detect(x_clean, "probably not") ~ 2,
    stringr::str_detect(x_clean, "probably yes") ~ 3,
    stringr::str_detect(x_clean, "definit") & stringr::str_detect(x_clean, "yes") ~ 4,
    TRUE ~ NA_real_
  )
}

# Create cleaned/re-coded IPSOS Youth object
IPSOS_Youth <- IPSOS_Youth_raw %>%
  dplyr::transmute(
    id = .data[[ipsos_id_col]],
    np_curious  = recode_ipsos_sus(NP13),
    np_nextyear = recode_ipsos_sus(NP11),
    np_soon     = recode_ipsos_sus(NP10),
    np_friend   = recode_ipsos_sus(NP12),
    source = "Ipsos youth",
    group = "Youth"
  )

# Check cleaned object
summary(IPSOS_Youth)
table(IPSOS_Youth$np_curious, useNA = "ifany")
table(IPSOS_Youth$np_nextyear, useNA = "ifany")
table(IPSOS_Youth$np_soon, useNA = "ifany")
table(IPSOS_Youth$np_friend, useNA = "ifany")

# Keep only IPSOS respondents with at least one pouch susceptibility item
IPSOS_Youth_analytic <- IPSOS_Youth[rowSums(!is.na(IPSOS_Youth[, c(
  "np_curious",
  "np_nextyear",
  "np_soon",
  "np_friend"
)])) > 0, ]

# Check analytic sample size
dim(IPSOS_Youth_analytic)

# Check distributions
summary(IPSOS_Youth_analytic)

```

#Create Pooled Analytic Sample
```{r Analytic sample}
# PATH pouch-only analytic file
PATH_np_analytic <- np_all_analytic %>%
  dplyr::select(
    np_curious, np_nextyear, np_soon, np_friend,
    source, group, dataset
  )

# IPSOS pouch-only analytic file
IPSOS_Youth_analytic <- IPSOS_Youth_analytic %>%
  dplyr::mutate(
    source = "Ipsos",
    dataset = "Ipsos youth",
    group = "Youth"
  ) %>%
  dplyr::select(
    np_curious, np_nextyear, np_soon, np_friend,
    source, group, dataset
  )

# Master pooled dataset
np_master <- dplyr::bind_rows(PATH_np_analytic, IPSOS_Youth_analytic)

# Remove objects from earlier workflow iterations
rm(
  da36498.8001,
  da36498.8002,
  IPSOS_Youth,
  list = ls(pattern = "^youth_np$|^ya_np$|^youth_np_recoded$|^ya_np_recoded$")
)
```

#Pouch Descriptives
```{r Item-level descriptives}
# Item means by dataset
aggregate(
  np_master[, c("np_curious", "np_nextyear", "np_soon", "np_friend")],
  by = list(dataset = np_master$dataset),
  FUN = function(x) mean(x, na.rm = TRUE)
)

# Item SDs by dataset
aggregate(
  np_master[, c("np_curious", "np_nextyear", "np_soon", "np_friend")],
  by = list(dataset = np_master$dataset),
  FUN = function(x) sd(x, na.rm = TRUE)
)

# Item means by source/group
aggregate(
  np_master[, c("np_curious", "np_nextyear", "np_soon", "np_friend")],
  by = list(source = np_master$source, group = np_master$group),
  FUN = function(x) mean(x, na.rm = TRUE)
)

# Response distributions by item and dataset
table(np_master$dataset, np_master$np_curious, useNA = "ifany")
table(np_master$dataset, np_master$np_nextyear, useNA = "ifany")
table(np_master$dataset, np_master$np_soon, useNA = "ifany")
table(np_master$dataset, np_master$np_friend, useNA = "ifany")


```
```{r Item-level correlations}
# Item names
np_items <- c("np_curious", "np_nextyear", "np_soon", "np_friend")

# Spearman correlations overall
cor(
  np_master[, np_items],
  use = "pairwise.complete.obs",
  method = "spearman"
)

# Spearman correlations by dataset
cor(
  np_master[np_master$dataset == "PATH youth", np_items],
  use = "pairwise.complete.obs",
  method = "spearman"
)

cor(
  np_master[np_master$dataset == "PATH young adult", np_items],
  use = "pairwise.complete.obs",
  method = "spearman"
)

cor(
  np_master[np_master$dataset == "Ipsos youth", np_items],
  use = "pairwise.complete.obs",
  method = "spearman"
)


```
```{r Reliability testing}
# Item names
np_item_names <- c("np_curious", "np_nextyear", "np_soon", "np_friend")

# Create clean numeric item dataset for reliability analyses
np_items <- np_master[, np_item_names]
np_items <- as.data.frame(lapply(np_items, as.numeric))

# Check structure
str(np_items)

# Overall alpha
psych::alpha(np_items)

# Alpha by dataset
psych::alpha(np_items[np_master$dataset == "PATH youth", ])
psych::alpha(np_items[np_master$dataset == "PATH young adult", ])
psych::alpha(np_items[np_master$dataset == "Ipsos youth", ])

# Overall omega
psych::omega(np_items, nfactors = 1)
```

#Pouch-specific CFA
```{r Pouch susceptibility CFA}
# Define one-factor model
np_model <- '
  np_sus =~ np_curious + np_nextyear + np_soon + np_friend
'

# Fit ordered categorical CFA
fit_np <- cfa(
  model = np_model,
  data = np_master,
  ordered = c("np_curious", "np_nextyear", "np_soon", "np_friend"),
  estimator = "WLSMV",
  missing = "pairwise"
)

# Model fit and standardized loadings
summary(fit_np, fit.measures = TRUE, standardized = TRUE)


```
```{r CFA by dataset}
# CFA among PATH youth
fit_np_path_youth <- lavaan::cfa(
  model = np_model,
  data = np_master[np_master$dataset == "PATH youth", ],
  ordered = c("np_curious", "np_nextyear", "np_soon", "np_friend"),
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_np_path_youth, fit.measures = TRUE, standardized = TRUE)

# CFA among PATH young adults
fit_np_path_ya <- lavaan::cfa(
  model = np_model,
  data = np_master[np_master$dataset == "PATH young adult", ],
  ordered = c("np_curious", "np_nextyear", "np_soon", "np_friend"),
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_np_path_ya, fit.measures = TRUE, standardized = TRUE)

# CFA among IPSOS youth
fit_np_ipsos_youth <- lavaan::cfa(
  model = np_model,
  data = np_master[np_master$dataset == "Ipsos youth", ],
  ordered = c("np_curious", "np_nextyear", "np_soon", "np_friend"),
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_np_ipsos_youth, fit.measures = TRUE, standardized = TRUE)
```
```{r Measurement invariance across age groups}
# Convert group to factor
np_master$group <- factor(np_master$group)

# Check group sizes
table(np_master$group, np_master$dataset)

# Configural invariance across age groups
fit_config_age <- lavaan::cfa(
  model = np_model,
  data = np_master,
  group = "group",
  ordered = c("np_curious", "np_nextyear", "np_soon", "np_friend"),
  estimator = "WLSMV",
  missing = "pairwise"
)

# Loadings + thresholds invariance for ordered items across age groups
fit_thresh_age <- lavaan::cfa(
  model = np_model,
  data = np_master,
  group = "group",
  ordered = c("np_curious", "np_nextyear", "np_soon", "np_friend"),
  estimator = "WLSMV",
  missing = "pairwise",
  group.equal = c("loadings", "thresholds")
)

# Compare fit
summary(fit_config_age, fit.measures = TRUE, standardized = TRUE)
summary(fit_thresh_age, fit.measures = TRUE, standardized = TRUE)

# Scaled chi-square difference test
lavaan::lavTestLRT(fit_config_age, fit_thresh_age)

# Extract practical fit measures
lavaan::fitMeasures(fit_config_age, c("cfi", "tli", "rmsea", "srmr"))
lavaan::fitMeasures(fit_thresh_age, c("cfi", "tli", "rmsea", "srmr"))

# Score test for equality constraints
lavaan::lavTestScore(fit_thresh_age)
```

#General vs product-specific CFA
```{r Product-specific CFA models}
# Run one-factor CFA for each product-specific susceptibility scale
product_cfa_results <- lapply(names(product_items), function(prod) {
  
  items <- product_items[[prod]]
  
  model_text <- paste0(
    prod, "_sus =~ ",
    paste(items, collapse = " + ")
  )
  
  fit <- lavaan::cfa(
    model = model_text,
    data = PATH_broad_sus_analytic,
    ordered = items,
    estimator = "WLSMV",
    missing = "pairwise"
  )
  
  list(
    product = prod,
    fit = fit,
    fit_measures = lavaan::fitMeasures(fit, c("cfi", "tli", "rmsea", "srmr"))
  )
})

# View fit measures
product_cfa_table <- dplyr::bind_rows(lapply(product_cfa_results, function(x) {
  data.frame(
    product = x$product,
    t(as.numeric(x$fit_measures))
  )
}))

names(product_cfa_table) <- c("product", "cfi", "tli", "rmsea", "srmr")

product_cfa_table
```
```{r One-factor CFA model}
# Single-factor model: all 36 ordered susceptibility items load on one factor
one_factor_model <- paste0(
  "general_sus =~ ",
  paste(broad_sus_items, collapse = " + ")
)

fit_one_factor <- lavaan::cfa(
  model = one_factor_model,
  data = PATH_broad_sus_analytic,
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_one_factor, fit.measures = TRUE, standardized = TRUE)

# Statistics reported in Table 2
one_factor_fit <- lavaan::fitMeasures(
  fit_one_factor,
  c("chisq.scaled", "df.scaled", "pvalue.scaled",
    "rmsea", "rmsea.ci.lower", "rmsea.ci.upper",
    "srmr", "cfi", "tli")
)
one_factor_fit
```
```{r Correlated product-specific model}
# Correlated product-specific factor model
product_factor_model <- '
  np_sus =~ np_curious + np_nextyear + np_soon + np_friend
  cig_sus =~ cig_curious + cig_nextyear + cig_soon + cig_friend
  ecig_sus =~ ecig_curious + ecig_nextyear + ecig_soon + ecig_friend
  hookah_sus =~ hookah_curious + hookah_nextyear + hookah_soon + hookah_friend
  snus_sus =~ snus_curious + snus_nextyear + snus_soon + snus_friend
  smkls_sus =~ smkls_curious + smkls_nextyear + smkls_soon + smkls_friend
  cigar_sus =~ cigar_curious + cigar_nextyear + cigar_soon + cigar_friend
  cigarillo_sus =~ cigarillo_curious + cigarillo_nextyear + cigarillo_soon + cigarillo_friend
  fcigar_sus =~ fcigar_curious + fcigar_nextyear + fcigar_soon + fcigar_friend
'

fit_product_factors <- lavaan::cfa(
  model = product_factor_model,
  data = PATH_broad_sus_analytic,
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_product_factors, fit.measures = TRUE, standardized = TRUE)

# Fit measures
lavaan::fitMeasures(
  fit_product_factors,
  c("cfi", "tli", "rmsea", "srmr")
)
```
```{r Product factor correlations}
# Extract standardized latent factor correlations
std_product <- lavaan::standardizedSolution(fit_product_factors)

product_factor_correlations <- std_product %>%
  dplyr::filter(op == "~~", lhs != rhs) %>%
  dplyr::select(lhs, rhs, est.std, pvalue) %>%
  dplyr::arrange(dplyr::desc(est.std))

product_factor_correlations

# Specifically inspect correlations involving nicotine pouch susceptibility
pouch_factor_correlations <- product_factor_correlations %>%
  dplyr::filter(lhs == "np_sus" | rhs == "np_sus") %>%
  dplyr::arrange(dplyr::desc(est.std))

pouch_factor_correlations
```
```{r Second-order general model}
general_factor_model <- '
  np_sus =~ np_curious + np_nextyear + np_soon + np_friend
  cig_sus =~ cig_curious + cig_nextyear + cig_soon + cig_friend
  ecig_sus =~ ecig_curious + ecig_nextyear + ecig_soon + ecig_friend
  hookah_sus =~ hookah_curious + hookah_nextyear + hookah_soon + hookah_friend
  snus_sus =~ snus_curious + snus_nextyear + snus_soon + snus_friend
  smkls_sus =~ smkls_curious + smkls_nextyear + smkls_soon + smkls_friend
  cigar_sus =~ cigar_curious + cigar_nextyear + cigar_soon + cigar_friend
  cigarillo_sus =~ cigarillo_curious + cigarillo_nextyear + cigarillo_soon + cigarillo_friend
  fcigar_sus =~ fcigar_curious + fcigar_nextyear + fcigar_soon + fcigar_friend

  general_sus =~ np_sus + cig_sus + ecig_sus + hookah_sus +
                 snus_sus + smkls_sus + cigar_sus + cigarillo_sus + fcigar_sus
'

fit_general_factor <- lavaan::cfa(
  model = general_factor_model,
  data = PATH_broad_sus_analytic,
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_general_factor, fit.measures = TRUE, standardized = TRUE)

lavaan::fitMeasures(
  fit_general_factor,
  c("cfi", "tli", "rmsea", "srmr"))
```
```{r Bifactor model}
bifactor_model <- '
  # General susceptibility factor
  general_sus =~ np_curious + np_nextyear + np_soon + np_friend +
                 cig_curious + cig_nextyear + cig_soon + cig_friend +
                 ecig_curious + ecig_nextyear + ecig_soon + ecig_friend +
                 hookah_curious + hookah_nextyear + hookah_soon + hookah_friend +
                 smkls_curious + smkls_nextyear + smkls_soon + smkls_friend +
                 snus_curious + snus_nextyear + snus_soon + snus_friend +
                 cigar_curious + cigar_nextyear + cigar_soon + cigar_friend +
                 cigarillo_curious + cigarillo_nextyear + cigarillo_soon + cigarillo_friend +
                 fcigar_curious + fcigar_nextyear + fcigar_soon + fcigar_friend

  # Product-specific residual factors
  np_specific =~ np_curious + np_nextyear + np_soon + np_friend
  cig_specific =~ cig_curious + cig_nextyear + cig_soon + cig_friend
  ecig_specific =~ ecig_curious + ecig_nextyear + ecig_soon + ecig_friend
  hookah_specific =~ hookah_curious + hookah_nextyear + hookah_soon + hookah_friend
  smkls_specific =~ smkls_curious + smkls_nextyear + smkls_soon + smkls_friend
  snus_specific =~ snus_curious + snus_nextyear + snus_soon + snus_friend
  cigar_specific =~ cigar_curious + cigar_nextyear + cigar_soon + cigar_friend
  cigarillo_specific =~ cigarillo_curious + cigarillo_nextyear + cigarillo_soon + cigarillo_friend
  fcigar_specific =~ fcigar_curious + fcigar_nextyear + fcigar_soon + fcigar_friend

  # Orthogonal factors
  general_sus ~~ 0*np_specific + 0*cig_specific + 0*ecig_specific +
                 0*hookah_specific + 0*smkls_specific + 0*snus_specific +
                 0*cigar_specific + 0*cigarillo_specific + 0*fcigar_specific

  np_specific ~~ 0*cig_specific + 0*ecig_specific + 0*hookah_specific +
                 0*smkls_specific + 0*snus_specific + 0*cigar_specific +
                 0*cigarillo_specific + 0*fcigar_specific

  cig_specific ~~ 0*ecig_specific + 0*hookah_specific + 0*smkls_specific +
                  0*snus_specific + 0*cigar_specific + 0*cigarillo_specific +
                  0*fcigar_specific

  ecig_specific ~~ 0*hookah_specific + 0*smkls_specific + 0*snus_specific +
                   0*cigar_specific + 0*cigarillo_specific + 0*fcigar_specific

  hookah_specific ~~ 0*smkls_specific + 0*snus_specific + 0*cigar_specific +
                     0*cigarillo_specific + 0*fcigar_specific

  smkls_specific ~~ 0*snus_specific + 0*cigar_specific +
                    0*cigarillo_specific + 0*fcigar_specific

  snus_specific ~~ 0*cigar_specific + 0*cigarillo_specific + 0*fcigar_specific

  cigar_specific ~~ 0*cigarillo_specific + 0*fcigar_specific

  cigarillo_specific ~~ 0*fcigar_specific
'

fit_bifactor <- lavaan::cfa(
  model = bifactor_model,
  data = PATH_broad_sus_analytic,
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_bifactor, fit.measures = TRUE, standardized = TRUE)

lavaan::fitMeasures(
  fit_bifactor,
  c("cfi", "tli", "rmsea", "srmr")
)
```
```{r Visual}
fit_df <- tibble::tribble(
  ~index, ~value, ~direction, ~good_cut, ~excellent_cut, ~label,
  "CFI",   .994, "higher", .95, .97, ".994",
  "TLI",   .993, "higher", .95, .97, ".993",
  "RMSEA", .050, "lower",  .08, .05, ".050",
  "SRMR",  .080, "lower",  .08, .05, ".080"
)

fit_df <- fit_df %>%
  mutate(
    status = case_when(
      direction == "higher" & value >= excellent_cut ~ "Excellent",
      direction == "higher" & value >= good_cut ~ "Good",
      direction == "lower" & value <= excellent_cut ~ "Excellent",
      direction == "lower" & value <= good_cut ~ "Good",
      TRUE ~ "Needs caution"
    ),
    index = factor(index, levels = rev(c("CFI", "TLI", "RMSEA", "SRMR")))
  )

ggplot(fit_df, aes(x = 1, y = index)) +
  geom_tile(aes(fill = status), width = .85, height = .7, color = "white") +
  geom_text(aes(label = paste0(index, " = ", label)), 
            size = 7, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c(
      "Excellent" = "#d9ead3",
      "Good" = "#fff2cc",
      "Needs caution" = "#f4cccc"
    )
  ) +
  labs(
    title = "Higher-order CFA model fit",
    subtitle = "The susceptibility measurement architecture fit the data well",
    fill = NULL
  ) +
  theme_void(base_size = 18) +
  theme(
    plot.title = element_text(face = "bold", size = 24),
    plot.subtitle = element_text(size = 16),
    legend.position = "bottom",
    plot.margin = margin(10, 20, 10, 20)
  )
```

#Measurement invariance
```{r Measurement invariance, eval=FALSE, purl=FALSE}
# Ensure group is a factor
PATH_broad_sus_analytic$group <- factor(PATH_broad_sus_analytic$group)

# Configural invariance: same factor structure across groups
fit_general_config <- lavaan::cfa(
  model = general_factor_model,
  data = PATH_broad_sus_analytic,
  group = "group",
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

# Threshold/loading invariance for ordered categorical items
fit_general_thresh <- lavaan::cfa(
  model = general_factor_model,
  data = PATH_broad_sus_analytic,
  group = "group",
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise",
  group.equal = c("loadings", "thresholds")
)

# Compare model fit
general_invariance_fit <- dplyr::bind_rows(
  data.frame(
    model = "Configural",
    t(lavaan::fitMeasures(
      fit_general_config,
      c("cfi", "tli", "rmsea", "srmr")
    ))
  ),
  data.frame(
    model = "Loadings + thresholds",
    t(lavaan::fitMeasures(
      fit_general_thresh,
      c("cfi", "tli", "rmsea", "srmr")
    ))
  )
)

general_invariance_fit <- general_invariance_fit %>%
  dplyr::mutate(
    dplyr::across(c(cfi, tli, rmsea, srmr), ~ round(.x, 3))
  )

general_invariance_fit

# Practical change in fit
general_invariance_delta <- general_invariance_fit %>%
  dplyr::summarise(
    delta_cfi = cfi[model == "Loadings + thresholds"] - cfi[model == "Configural"],
    delta_tli = tli[model == "Loadings + thresholds"] - tli[model == "Configural"],
    delta_rmsea = rmsea[model == "Loadings + thresholds"] - rmsea[model == "Configural"],
    delta_srmr = srmr[model == "Loadings + thresholds"] - srmr[model == "Configural"]
  )

general_invariance_delta

# Chi-square difference test - expected to be sensitive with this sample size
lavaan::lavTestLRT(fit_general_config, fit_general_thresh)
```
```{r Full-scale invariance recode sparse categories}
PATH_broad_inv <- PATH_broad_sus_analytic

PATH_broad_inv[broad_sus_items] <- lapply(
  PATH_broad_inv[broad_sus_items],
  function(x) {
    x_num <- as.numeric(x)

    dplyr::case_when(
      is.na(x_num) ~ NA_real_,
      x_num == 1 ~ 1,
      x_num == 2 ~ 2,
      x_num %in% c(3, 4) ~ 3
    )
  }
)

# Quick check for empty categories by group
empty_category_check <- lapply(broad_sus_items, function(v) {
  tab <- table(PATH_broad_inv$group, PATH_broad_inv[[v]], useNA = "no")
  data.frame(
    item = v,
    has_empty_category = any(tab == 0)
  )
}) %>%
  dplyr::bind_rows()

empty_category_check %>%
  dplyr::filter(has_empty_category)
```
```{r Full-scale measurement invariance collapsed}
PATH_broad_inv$group <- factor(PATH_broad_inv$group)

fit_general_config_collapsed <- lavaan::cfa(
  model = general_factor_model,
  data = PATH_broad_inv,
  group = "group",
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

fit_general_thresh_collapsed <- lavaan::cfa(
  model = general_factor_model,
  data = PATH_broad_inv,
  group = "group",
  ordered = broad_sus_items,
  estimator = "WLSMV",
  missing = "pairwise",
  group.equal = c("loadings", "thresholds")
)

# Helper to extract fit measures as a clean one-row data frame
get_lavaan_fit <- function(fit, model_name) {
  fit_values <- lavaan::fitMeasures(
    fit,
    c("cfi", "tli", "rmsea", "srmr")
  )

  data.frame(
    model = model_name,
    cfi = as.numeric(fit_values["cfi"]),
    tli = as.numeric(fit_values["tli"]),
    rmsea = as.numeric(fit_values["rmsea"]),
    srmr = as.numeric(fit_values["srmr"])
  )
}

general_invariance_fit_collapsed <- dplyr::bind_rows(
  get_lavaan_fit(fit_general_config_collapsed, "Configural"),
  get_lavaan_fit(fit_general_thresh_collapsed, "Loadings + thresholds")
) %>%
  dplyr::mutate(
    dplyr::across(c(cfi, tli, rmsea, srmr), ~ round(.x, 3))
  )

general_invariance_fit_collapsed

general_invariance_delta_collapsed <- general_invariance_fit_collapsed %>%
  dplyr::summarise(
    delta_cfi = cfi[model == "Loadings + thresholds"] - cfi[model == "Configural"],
    delta_tli = tli[model == "Loadings + thresholds"] - tli[model == "Configural"],
    delta_rmsea = rmsea[model == "Loadings + thresholds"] - rmsea[model == "Configural"],
    delta_srmr = srmr[model == "Loadings + thresholds"] - srmr[model == "Configural"]
  )

general_invariance_delta_collapsed
```
```{r Full-scale measurement invariance collapsed fit table}
# Helper to extract fit measures as a clean one-row data frame
get_lavaan_fit <- function(fit, model_name) {
  fit_values <- lavaan::fitMeasures(
    fit,
    c("cfi", "tli", "rmsea", "srmr")
  )

  data.frame(
    model = model_name,
    cfi = as.numeric(fit_values["cfi"]),
    tli = as.numeric(fit_values["tli"]),
    rmsea = as.numeric(fit_values["rmsea"]),
    srmr = as.numeric(fit_values["srmr"])
  )
}

general_invariance_fit_collapsed <- dplyr::bind_rows(
  get_lavaan_fit(fit_general_config_collapsed, "Configural"),
  get_lavaan_fit(fit_general_thresh_collapsed, "Loadings + thresholds")
) %>%
  dplyr::mutate(
    dplyr::across(c(cfi, tli, rmsea, srmr), ~ round(.x, 3))
  )

general_invariance_fit_collapsed

general_invariance_delta_collapsed <- general_invariance_fit_collapsed %>%
  dplyr::summarise(
    delta_cfi = cfi[model == "Loadings + thresholds"] - cfi[model == "Configural"],
    delta_tli = tli[model == "Loadings + thresholds"] - tli[model == "Configural"],
    delta_rmsea = rmsea[model == "Loadings + thresholds"] - rmsea[model == "Configural"],
    delta_srmr = srmr[model == "Loadings + thresholds"] - srmr[model == "Configural"]
  )

general_invariance_delta_collapsed
```

#Pouch-specific IRT
```{r Pouch IRT data}
# Item names
np_item_names <- c("np_curious", "np_nextyear", "np_soon", "np_friend")

# Create IRT dataset from pooled master sample
NP_IRT <- np_master %>%
  dplyr::select(dplyr::all_of(np_item_names)) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.numeric))

# Apply listwise deletion for IRT analyses
NP_IRT <- NP_IRT[complete.cases(NP_IRT), ]

# Check final IRT sample
summary(NP_IRT)
dim(NP_IRT)


```
```{r Mokken scale analysis}
# Mokken nonparametric IRT
NP_IRT_mokken <- NP_IRT %>%
  mutate(across(everything(), as.numeric)) %>%
  as.data.frame()

# Scalability
coefH(NP_IRT_mokken)

# Monotonicity
NP_monotonicity <- check.monotonicity(NP_IRT_mokken)
summary(NP_monotonicity)
plot(NP_monotonicity)

# Reliability
check.reliability(NP_IRT_mokken)
```
```{r Graded response model}
# Fit 1-factor graded response model
fit_NP_grm <- mirt(
  data = NP_IRT,
  model = 1,
  itemtype = "graded",
  technical = list(removeEmptyRows = TRUE),
  SE = TRUE
)

summary(fit_NP_grm)
```
```{r Item parameters}
# IRT parameters
NP_coef <- coef(fit_NP_grm, IRTpars = TRUE, simplify = TRUE)

# Global fit
NP_M2 <- tryCatch(
  mirt::M2(fit_NP_grm, type = "M2*"),
  error = function(e) data.frame(note = conditionMessage(e))
)
NP_M2

# Residual correlation matrix for local dependence
resid_results <- residuals(fit_NP_grm, type = "Q3")
resid_results
```
```{r Item parameters table}
# Convert GRM item parameters to data frame
grm_params <- as.data.frame(NP_coef$items) %>%
  tibble::rownames_to_column(var = "Variable")

# Item labels in harmonized order
np_labels <- c(
  "Curious about using nicotine pouches",
  "Use nicotine pouches in the next year",
  "Try nicotine pouches soon",
  "Use nicotine pouches if best friend offered"
)

# Build clean GRM item parameter table
table2_grm <- grm_params %>%
  dplyr::mutate(
    Item_ID = c("NP_Curious", "NP_NextYear", "NP_Soon", "NP_Friend"),
    Item = paste0(np_labels, " (", Item_ID, ")")
  ) %>%
  dplyr::select(Item, a, b1, b2, b3) %>%
  dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 2)))

# View table
table2_grm

# Export table
write.csv(table2_grm, "Table2_GRM_item_parameters_NP.csv", row.names = FALSE)




```
```{r Option and item characeristic curves}
plot(fit_NP_grm, type = "trace", facet_items = TRUE)
plot(fit_NP_grm, type = "infotrace", facet_items = TRUE)

```
```{r Test information function}
theta_grid <- seq(-3, 3, length.out = 500)

test_info <- mirt::testinfo(
  fit_NP_grm,
  Theta = matrix(theta_grid, ncol = 1)
)

test_se <- 1 / sqrt(test_info)

NP_info_se_df <- data.frame(
  theta = theta_grid,
  test_info = test_info,
  se = test_se
)

# Scale SE to information range for plotting
scale_factor <- max(NP_info_se_df$test_info, na.rm = TRUE) /
  max(NP_info_se_df$se, na.rm = TRUE)

ggplot(NP_info_se_df, aes(x = theta)) +
  geom_line(aes(y = test_info, color = "Test Information"), linewidth = 1.2) +
  geom_line(aes(y = se * scale_factor, color = "Standard Error"),
            linewidth = 1.2, linetype = "dashed") +
  scale_y_continuous(
    name = "Test Information",
    sec.axis = sec_axis(~ . / scale_factor, name = "Standard Error")
  ) +
  scale_color_manual(
    values = c("Test Information" = "black", "Standard Error" = "gray40")
  ) +
  theme_bw() +
  labs(
    x = expression(theta),
    color = NULL,
    title = "Test Information and Standard Error"
  )
```
```{r Conditional reliability}
marginal_rxx(fit_NP_grm)

theta_grid <- seq(-3, 3, length.out = 500)

test_info <- mirt::testinfo(
  fit_NP_grm,
  Theta = matrix(theta_grid, ncol = 1)
)

cond_rel <- test_info / (test_info + 1)

plot(
  theta_grid,
  cond_rel,
  type = "l",
  lwd = 3,
  col = "blue",
  ylim = c(0, 1),
  xlab = expression(theta),
  ylab = "Conditional Reliability",
  main = "Conditional Reliability Curve"
)

abline(h = c(.70, .80, .90), lty = 2, col = "gray")
```
```{r Item information curves}
# Item information curves
item_info_list <- lapply(seq_along(np_item_names), function(i) {
  data.frame(
    theta = as.vector(theta_grid),
    info  = mirt::iteminfo(mirt::extract.item(fit_NP_grm, i), Theta = theta_grid),
    item  = np_item_names[i]
  )
})

item_info_df <- dplyr::bind_rows(item_info_list)

ggplot(item_info_df, aes(x = theta, y = info)) +
  geom_line(linewidth = 1) +
  facet_wrap(~ item, scales = "free_y") +
  labs(
    title = "Item Information Curves",
    x = expression(theta),
    y = "Item Information"
  ) +
  theme_minimal(base_size = 13)


```
```{r Differential item functioning}
# Create complete-case dataset for IRT DIF
NP_DIF <- np_master %>%
  dplyr::select(group, dplyr::all_of(np_item_names)) %>%
  dplyr::mutate(
    group = factor(group),
    dplyr::across(dplyr::all_of(np_item_names), as.numeric)
  ) %>%
  tidyr::drop_na(dplyr::all_of(np_item_names), group)

# Check group sizes
table(NP_DIF$group)

# Item matrix and group vector
NP_DIF_items <- NP_DIF %>%
  dplyr::select(dplyr::all_of(np_item_names))

NP_DIF_group <- NP_DIF$group

# Multiple-group graded response model
fit_NP_mg <- mirt::multipleGroup(
  data = NP_DIF_items,
  model = 1,
  group = NP_DIF_group,
  itemtype = "graded",
  invariance = c("slopes", "intercepts", "free_means", "free_var")
)

summary(fit_NP_mg)

DIF_NP_age <- mirt::DIF(
  fit_NP_mg,
  which.par = c("a1", "d1", "d2", "d3"),
  scheme = "drop",
  items2test = 1:length(np_item_names),
  p.adjust = "BH"
)

DIF_NP_age

# Fit less-constrained multiple-group model for inspecting group-specific item parameters
# Anchor first item only to identify the scale
fit_NP_mg_free <- mirt::multipleGroup(
  data = NP_DIF_items,
  model = 1,
  group = NP_DIF_group,
  itemtype = "graded",
  invariance = c("np_curious", "free_means", "free_var")
)

summary(fit_NP_mg_free)

# Extract group-specific IRT parameters
coef_mg_free <- mirt::coef(fit_NP_mg_free, IRTpars = TRUE, simplify = TRUE)

coef_mg_free$Youth$items
coef_mg_free$`Young adult`$items

```

#Product-specific IRT
```{r Product comparator IRT setup}
# Product labels for cleaner tables
product_labels <- c(
  np = "Nicotine pouches",
  cig = "Cigarettes",
  ecig = "E-cigarettes",
  hookah = "Hookah",
  snus = "Snus",
  smkls = "Smokeless tobacco",
  cigar = "Traditional cigars",
  cigarillo = "Cigarillos",
  fcigar = "Filtered cigars"
)

# Item labels shared across product-specific susceptibility scales
sus_item_labels <- c(
  curious = "Curious",
  nextyear = "Use in next year",
  soon = "Try soon",
  friend = "Best friend offered"
)
```
```{r Product comparator IRT product universe}
# Product universe for comparator IRT analyses
product_items_main <- product_items[c(
  "np",
  "cig",
  "ecig",
  "hookah",
  "smkls",
  "snus",
  "cigar",
  "cigarillo",
  "fcigar"
)]

broad_sus_items_main <- unname(unlist(product_items_main))

# Check that all selected variables exist
setdiff(broad_sus_items_main, names(PATH_broad_sus_analytic))
```
```{r Product comparator Mokken}
product_mokken_results <- lapply(names(product_items_main), function(prod) {
  
  items <- product_items_main[[prod]]
  
  dat <- PATH_broad_sus_analytic %>%
    dplyr::select(dplyr::all_of(items)) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), as.numeric)) %>%
    tidyr::drop_na()
  
  if (nrow(dat) < 100) return(NULL)
  
  h <- mokken::coefH(dat)
  
  data.frame(
    product = prod,
    product_label = product_labels[prod],
    n = nrow(dat),
    scale_H = as.numeric(h$H[1]),
    stringsAsFactors = FALSE
  )
})

product_mokken_table <- dplyr::bind_rows(product_mokken_results) %>%
  dplyr::mutate(scale_H = round(scale_H, 3))

product_mokken_table
```
```{r Product comparator GRM}
fit_product_grm <- lapply(names(product_items_main), function(prod) {
  
  items <- product_items_main[[prod]]
  
  dat <- PATH_broad_sus_analytic %>%
    dplyr::select(dplyr::all_of(items)) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), as.numeric)) %>%
    tidyr::drop_na()
  
  if (nrow(dat) < 100) return(NULL)
  
  fit <- mirt::mirt(
    data = dat,
    model = 1,
    itemtype = "graded",
    verbose = FALSE
  )
  
  list(
    product = prod,
    product_label = unname(product_labels[prod]),
    n = nrow(dat),
    fit = fit
  )
})

names(fit_product_grm) <- names(product_items_main)
fit_product_grm <- fit_product_grm[!sapply(fit_product_grm, is.null)]
```
```{r Product comparator item parameters}
# Extract GRM item parameters for each product
product_grm_params <- lapply(fit_product_grm, function(x) {
  
  params <- mirt::coef(
    x$fit,
    IRTpars = TRUE,
    simplify = TRUE
  )$items %>%
    as.data.frame() %>%
    tibble::rownames_to_column("item_name")
  
  params %>%
    dplyr::mutate(
      product = x$product,
      product_label = x$product_label,
      n = x$n,
      item_type = dplyr::case_when(
        stringr::str_detect(item_name, "curious") ~ "Curious",
        stringr::str_detect(item_name, "nextyear") ~ "Use in next year",
        stringr::str_detect(item_name, "soon") ~ "Try soon",
        stringr::str_detect(item_name, "friend") ~ "Best friend offered",
        TRUE ~ item_name
      )
    ) %>%
    dplyr::select(
      product,
      product_label,
      n,
      item_name,
      item_type,
      a,
      b1,
      b2,
      b3
    )
})

product_grm_param_table <- dplyr::bind_rows(product_grm_params) %>%
  dplyr::mutate(
    dplyr::across(c(a, b1, b2, b3), ~ round(.x, 2))
  )

product_grm_param_table
```

#IRT visuals
```{r Product IRT reporting tables}
# Order item types for clean tables/figures
item_type_order <- c(
  "Curious",
  "Use in next year",
  "Try soon",
  "Best friend offered"
)

# Clean GRM parameter table
product_grm_report_table <- product_grm_param_table %>%
  dplyr::mutate(
    item_type = factor(item_type, levels = item_type_order),
    product_label = factor(product_label, levels = unique(product_label))
  ) %>%
  dplyr::arrange(product_label, item_type) %>%
  dplyr::select(
    product_label,
    n,
    item_type,
    a,
    b1,
    b2,
    b3
  )

product_grm_report_table

# Most and second-most discriminating item by product
product_discrimination_rank <- product_grm_param_table %>%
  dplyr::group_by(product, product_label) %>%
  dplyr::arrange(dplyr::desc(a), .by_group = TRUE) %>%
  dplyr::mutate(rank = dplyr::row_number()) %>%
  dplyr::ungroup()

top_discrimination_table <- product_discrimination_rank %>%
  dplyr::filter(rank <= 2) %>%
  dplyr::select(product_label, rank, item_type, a) %>%
  tidyr::pivot_wider(
    names_from = rank,
    values_from = c(item_type, a),
    names_glue = "rank_{rank}_{.value}"
  ) %>%
  dplyr::rename(
    most_discriminating_item = rank_1_item_type,
    most_discriminating_a = rank_1_a,
    second_discriminating_item = rank_2_item_type,
    second_discriminating_a = rank_2_a
  ) %>%
  dplyr::mutate(
    most_discriminating_a = round(most_discriminating_a, 2),
    second_discriminating_a = round(second_discriminating_a, 2)
  )

top_discrimination_table

# Quick check: which item wins by product?
table(top_discrimination_table$most_discriminating_item)
```
```{r Product IRT discrimination heatmap}
item_type_order <- c(
  "Curious",
  "Best friend offered",
  "Try soon",
  "Use in next year"
)

# Order products by discrimination for "Use in next year"
product_order_nextyear <- product_grm_param_table %>%
  dplyr::filter(item_type == "Use in next year") %>%
  dplyr::arrange(dplyr::desc(a)) %>%
  dplyr::pull(product_label)

product_grm_param_table %>%
  dplyr::mutate(
    item_type = factor(item_type, levels = item_type_order),
    product_label = factor(product_label, levels = rev(product_order_nextyear))
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = item_type,
      y = product_label,
      fill = a
    )
  ) +
  ggplot2::geom_tile(color = "white", linewidth = 0.6) +
  ggplot2::geom_text(
    ggplot2::aes(label = round(a, 2)),
    size = 3.5
  ) +
  ggplot2::scale_fill_gradient(
    low = "white",
    high = "steelblue",
    name = "Discrimination\n(a)"
  ) +
  ggplot2::theme_bw(base_size = 13) +
  ggplot2::theme(
    axis.title = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank(),
    legend.position = "right"
  ) +
  ggplot2::labs(
    title = "IRT Discrimination Parameters by Product and Item"
  )
```
```{r General susceptibility IRT information}
# Create item matrix for broad/general susceptibility IRT
general_irt_items <- PATH_broad_sus_analytic %>%
  dplyr::select(dplyr::all_of(broad_sus_items)) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.numeric))

# Keep rows with at least one susceptibility item answered
general_irt_items <- general_irt_items[
  rowSums(!is.na(general_irt_items)) > 0,
]

# Fit one-factor GRM across entire product universe
fit_general_grm <- mirt::mirt(
  data = general_irt_items,
  model = 1,
  itemtype = "graded",
  verbose = TRUE
)

summary(fit_general_grm)

# Test information and standard error
theta_grid <- seq(-3, 4, by = 0.05)

general_test_info <- mirt::testinfo(
  fit_general_grm,
  Theta = matrix(theta_grid)
)

general_test_se <- 1 / sqrt(general_test_info)

general_info_df <- data.frame(
  theta = theta_grid,
  test_info = general_test_info,
  se = general_test_se
)

# Plot test information and SE together
scale_factor <- max(general_info_df$test_info, na.rm = TRUE) /
  max(general_info_df$se, na.rm = TRUE)

ggplot2::ggplot(general_info_df, ggplot2::aes(x = theta)) +
  ggplot2::geom_line(
    ggplot2::aes(y = test_info, linetype = "Test information"),
    linewidth = 1.1,
    color = "black"
  ) +
  ggplot2::geom_line(
    ggplot2::aes(y = se * scale_factor, linetype = "Standard error"),
    linewidth = 1.1,
    color = "gray40"
  ) +
  ggplot2::scale_y_continuous(
    name = "Test Information",
    sec.axis = ggplot2::sec_axis(
      ~ . / scale_factor,
      name = "Standard Error"
    )
  ) +
  ggplot2::scale_linetype_manual(
    values = c(
      "Test information" = "solid",
      "Standard error" = "dashed"
    ),
    name = NULL
  ) +
  ggplot2::theme_bw(base_size = 13) +
  ggplot2::theme(
    legend.position = "right"
  ) +
  ggplot2::labs(
    title = "General Tobacco Susceptibility Test Information",
    x = expression(theta)
  )
```
```{r General susceptibility conditional reliability}
general_cond_rel <- general_test_info / (general_test_info + 1)

general_rel_df <- data.frame(
  theta = theta_grid,
  conditional_reliability = general_cond_rel
)

ggplot2::ggplot(
  general_rel_df,
  ggplot2::aes(x = theta, y = conditional_reliability)
) +
  ggplot2::geom_line(linewidth = 1.1, color = "black") +
  ggplot2::geom_hline(
    yintercept = c(.70, .80, .90),
    linetype = "dashed",
    color = "gray70"
  ) +
  ggplot2::coord_cartesian(ylim = c(0, 1)) +
  ggplot2::theme_bw(base_size = 13) +
  ggplot2::labs(
    title = "Conditional Reliability for General Tobacco Susceptibility",
    x = expression(theta),
    y = "Conditional Reliability"
  )
```
```{r Next-year maximum information by product}
theta_grid <- matrix(seq(-3, 4, by = 0.05))

nextyear_info_list <- lapply(fit_product_grm, function(x) {
  item_names <- rownames(
    mirt::coef(x$fit, IRTpars = TRUE, simplify = TRUE)$items
  )

  nextyear_position <- which(stringr::str_detect(item_names, "nextyear"))

  info <- mirt::iteminfo(
    mirt::extract.item(x$fit, nextyear_position),
    Theta = theta_grid
  )

  data.frame(
    theta = as.vector(theta_grid),
    information = info,
    product = x$product,
    product_label = x$product_label
  )
})

nextyear_info_df <- dplyr::bind_rows(nextyear_info_list)

nextyear_info_summary <- nextyear_info_df %>%
  dplyr::group_by(product, product_label) %>%
  dplyr::summarise(
    max_information = max(information, na.rm = TRUE),
    theta_at_max_information = theta[which.max(information)],
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    max_information = round(max_information, 2),
    theta_at_max_information = round(theta_at_max_information, 2)
  ) %>%
  dplyr::arrange(dplyr::desc(max_information))

nextyear_info_summary

ggplot2::ggplot(
  nextyear_info_summary,
  ggplot2::aes(
    x = reorder(product_label, max_information),
    y = max_information
  )
) +
  ggplot2::geom_point(size = 3) +
  ggplot2::coord_flip() +
  ggplot2::theme_bw(base_size = 13) +
  ggplot2::labs(
    title = "Maximum Information for Next-Year Intention Item",
    x = NULL,
    y = "Maximum Item Information"
  )
```

#Short-form screener
```{r Short-form susceptibility setup}
# Full product universe, IQOS excluded
product_names_main <- names(product_items)

nextyear_items <- paste0(product_names_main, "_nextyear")
soon_items <- paste0(product_names_main, "_soon")

short_nextyear_items <- nextyear_items
short_intention_items <- c(nextyear_items, soon_items)

# Check variables exist
setdiff(short_nextyear_items, names(PATH_broad_sus_analytic))
setdiff(short_intention_items, names(PATH_broad_sus_analytic))
```
```{r Short-form CFA models}
# Helper to extract fit measures as a clean one-row data frame
get_lavaan_fit <- function(fit, model_name) {
  fit_values <- lavaan::fitMeasures(
    fit,
    c("cfi", "tli", "rmsea", "srmr")
  )

  data.frame(
    model = model_name,
    cfi = as.numeric(fit_values["cfi"]),
    tli = as.numeric(fit_values["tli"]),
    rmsea = as.numeric(fit_values["rmsea"]),
    srmr = as.numeric(fit_values["srmr"])
  )
}

# Next-year only: one general susceptibility factor
nextyear_model <- paste0(
  "general_sus =~ ",
  paste(short_nextyear_items, collapse = " + ")
)

nextyear_data <- PATH_broad_sus_analytic %>%
  dplyr::filter(rowSums(!is.na(dplyr::select(., dplyr::all_of(short_nextyear_items)))) > 0)

fit_nextyear <- lavaan::cfa(
  model = nextyear_model,
  data = nextyear_data,
  ordered = short_nextyear_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

# Try-soon only: one general susceptibility factor
trysoon_items <- paste0(product_names_main, "_soon")

trysoon_model <- paste0(
  "general_sus =~ ",
  paste(trysoon_items, collapse = " + ")
)

trysoon_data <- PATH_broad_sus_analytic %>%
  dplyr::filter(rowSums(!is.na(dplyr::select(., dplyr::all_of(trysoon_items)))) > 0)

fit_trysoon <- lavaan::cfa(
  model = trysoon_model,
  data = trysoon_data,
  ordered = trysoon_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

# Next-year + try-soon: one general susceptibility factor
intention_model <- paste0(
  "general_sus =~ ",
  paste(short_intention_items, collapse = " + ")
)

intention_data <- PATH_broad_sus_analytic %>%
  dplyr::filter(rowSums(!is.na(dplyr::select(., dplyr::all_of(short_intention_items)))) > 0)

fit_intention <- lavaan::cfa(
  model = intention_model,
  data = intention_data,
  ordered = short_intention_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

# Summaries if needed
summary(fit_nextyear, fit.measures = TRUE, standardized = TRUE)
summary(fit_trysoon, fit.measures = TRUE, standardized = TRUE)
summary(fit_intention, fit.measures = TRUE, standardized = TRUE)

# Short-form model fit comparison
short_form_fit <- dplyr::bind_rows(
  get_lavaan_fit(fit_general_factor, "Full 36-item second-order model"),
  get_lavaan_fit(fit_nextyear, "Short form: next-year only"),
  get_lavaan_fit(fit_trysoon, "Short form: try-soon only"),
  get_lavaan_fit(fit_intention, "Short form: next-year + try-soon")
) %>%
  dplyr::mutate(
    dplyr::across(c(cfi, tli, rmsea, srmr), ~ round(.x, 3))
  )

short_form_fit
```
```{r Short-form IRT information}
# Helper to fit general GRM and extract information curve
fit_short_grm_info <- function(data, items, label, theta_grid = seq(-3, 4, by = 0.05)) {
  item_data <- data %>%
    dplyr::select(dplyr::all_of(items)) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), as.numeric))

  item_data <- item_data[rowSums(!is.na(item_data)) > 0, ]

  fit <- mirt::mirt(
    data = item_data,
    model = 1,
    itemtype = "graded",
    verbose = FALSE
  )

  info <- mirt::testinfo(fit, Theta = matrix(theta_grid))

  list(
    fit = fit,
    info = data.frame(
      theta = theta_grid,
      test_info = info,
      form = label
    )
  )
}

full_grm_info <- fit_short_grm_info(
  PATH_broad_sus_analytic,
  broad_sus_items,
  "Full 36-item form"
)

nextyear_grm_info <- fit_short_grm_info(
  PATH_broad_sus_analytic,
  short_nextyear_items,
  "Next-year only"
)

intention_grm_info <- fit_short_grm_info(
  PATH_broad_sus_analytic,
  short_intention_items,
  "Next-year + try-soon"
)

short_form_info_df <- dplyr::bind_rows(
  full_grm_info$info,
  nextyear_grm_info$info,
  intention_grm_info$info
)

ggplot2::ggplot(
  short_form_info_df,
  ggplot2::aes(x = theta, y = test_info, color = form, linetype = form)
) +
  ggplot2::geom_line(linewidth = 1.1) +
  ggplot2::theme_bw(base_size = 13) +
  ggplot2::labs(
    title = "Test Information by Susceptibility Battery Length",
    x = expression(theta),
    y = "Test Information",
    color = NULL,
    linetype = NULL
  )
```
```{r Short-form information summary}
short_form_info_summary <- short_form_info_df %>%
  dplyr::group_by(form) %>%
  dplyr::summarise(
    max_information = max(test_info, na.rm = TRUE),
    theta_at_max_information = theta[which.max(test_info)],
    mean_information_theta_0_to_3 = mean(test_info[theta >= 0 & theta <= 3], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    dplyr::across(where(is.numeric), ~ round(.x, 2))
  ) %>%
  dplyr::arrange(dplyr::desc(max_information))

short_form_info_summary
```
```{r Short-form try-soon only}
trysoon_items <- paste0(product_names_main, "_soon")

# Check variables exist
setdiff(trysoon_items, names(PATH_broad_sus_analytic))

trysoon_model <- paste0(
  "general_sus =~ ",
  paste(trysoon_items, collapse = " + ")
)

fit_trysoon <- lavaan::cfa(
  model = trysoon_model,
  data = PATH_broad_sus_analytic,
  ordered = trysoon_items,
  estimator = "WLSMV",
  missing = "pairwise"
)

summary(fit_trysoon, fit.measures = TRUE, standardized = TRUE)

# Updated short-form fit table
short_form_fit <- dplyr::bind_rows(
  get_lavaan_fit(fit_general_factor, "Full 36-item second-order model"),
  get_lavaan_fit(fit_nextyear, "Short form: next-year only"),
  get_lavaan_fit(fit_trysoon, "Short form: try-soon only"),
  get_lavaan_fit(fit_intention, "Short form: next-year + try-soon")
) %>%
  dplyr::mutate(
    dplyr::across(c(cfi, tli, rmsea, srmr), ~ round(.x, 3))
  )

short_form_fit
```
```{r Short-form IRT information with try-soon}
trysoon_grm_info <- fit_short_grm_info(
  PATH_broad_sus_analytic,
  trysoon_items,
  "Try-soon only"
)

short_form_info_df <- dplyr::bind_rows(
  full_grm_info$info,
  nextyear_grm_info$info,
  trysoon_grm_info$info,
  intention_grm_info$info
)

ggplot2::ggplot(
  short_form_info_df,
  ggplot2::aes(x = theta, y = test_info, color = form, linetype = form)
) +
  ggplot2::geom_line(linewidth = 1.1) +
  ggplot2::theme_bw(base_size = 13) +
  ggplot2::labs(
    title = "Test Information by Susceptibility Battery Length",
    x = expression(theta),
    y = "Test Information",
    color = NULL,
    linetype = NULL
  )

short_form_info_summary <- short_form_info_df %>%
  dplyr::group_by(form) %>%
  dplyr::summarise(
    max_information = max(test_info, na.rm = TRUE),
    theta_at_max_information = theta[which.max(test_info)],
    mean_information_theta_0_to_3 = mean(test_info[theta >= 0 & theta <= 3], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    dplyr::across(where(is.numeric), ~ round(.x, 2))
  ) %>%
  dplyr::arrange(dplyr::desc(max_information))

short_form_info_summary
```
```{r Try-soon OCC by product}
# Theta grid
theta_grid <- matrix(seq(-3, 4, by = 0.05))

# Extract OCCs for try-soon item from each product-specific GRM
trysoon_occ_list <- lapply(fit_product_grm, function(x) {
  item_names <- rownames(
    mirt::coef(x$fit, IRTpars = TRUE, simplify = TRUE)$items
  )

  trysoon_position <- which(stringr::str_detect(item_names, "_soon$"))

  probs <- mirt::probtrace(
    mirt::extract.item(x$fit, trysoon_position),
    Theta = theta_grid
  )

  probs_df <- as.data.frame(probs)
  colnames(probs_df) <- paste0("Response_", seq_len(ncol(probs_df)))

  probs_df %>%
    dplyr::mutate(
      theta = as.vector(theta_grid),
      product = x$product,
      product_label = x$product_label
    ) %>%
    tidyr::pivot_longer(
      cols = dplyr::starts_with("Response_"),
      names_to = "response",
      values_to = "probability"
    ) %>%
    dplyr::mutate(
      response = dplyr::case_when(
        response == "Response_1" ~ "Definitely not",
        response == "Response_2" ~ "Probably not",
        response == "Response_3" ~ "Probably yes",
        response == "Response_4" ~ "Definitely yes",
        TRUE ~ response
      ),
      response = factor(
        response,
        levels = c(
          "Definitely not",
          "Probably not",
          "Probably yes",
          "Definitely yes"
        )
      )
    )
})

trysoon_occ_df <- dplyr::bind_rows(trysoon_occ_list)

# Plot OCCs by product
ggplot2::ggplot(
  trysoon_occ_df,
  ggplot2::aes(
    x = theta,
    y = probability,
    color = response,
    linetype = response
  )
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::facet_wrap(~ product_label, ncol = 3) +
  ggplot2::coord_cartesian(xlim = c(-1, 4), ylim = c(0, 1)) +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::theme(
    legend.position = "bottom",
    panel.grid.minor = ggplot2::element_blank()
  ) +
  ggplot2::labs(
    title = "Option Characteristic Curves for Try-Soon Items",
    x = expression(theta),
    y = "Response Probability",
    color = NULL,
    linetype = NULL
  )
```
#Practical scoring and revised information figures

```{r Product-specific observed subscale scores}
# Create practical product scores on the original 1-4 response metric.
# A product score is calculated only when all four constituent items are observed.
product_score_data <- PATH_broad_sus_analytic

for (prod in names(product_items_main)) {
  items <- product_items_main[[prod]]
  score_name <- paste0(prod, "_score")

  product_score_data[[score_name]] <- ifelse(
    rowSums(!is.na(product_score_data[, items])) == length(items),
    rowMeans(product_score_data[, items], na.rm = FALSE),
    NA_real_
  )
}

product_score_names <- paste0(names(product_items_main), "_score")

product_score_data$n_product_scores <- rowSums(
  !is.na(product_score_data[, product_score_names])
)

# The availability distribution is essential because PATH routing differs by product.
product_score_availability <- product_score_data %>%
  dplyr::count(n_product_scores, name = "n") %>%
  dplyr::mutate(percent = 100 * n / sum(n))

product_score_availability
```

```{r Product-specific ordinal omega}
# Estimate categorical omega separately for each four-item product scale.
# Complete cases are used within product to match the product-specific GRM samples.
product_reliability_results <- lapply(names(product_items_main), function(prod) {
  items <- product_items_main[[prod]]

  dat <- PATH_broad_sus_analytic %>%
    dplyr::select(dplyr::all_of(items)) %>%
    tidyr::drop_na()

  model_text <- paste0(
    prod, "_sus =~ ",
    paste(items, collapse = " + ")
  )

  fit <- lavaan::cfa(
    model = model_text,
    data = dat,
    ordered = items,
    estimator = "WLSMV"
  )

  data.frame(
    product = prod,
    product_label = unname(product_labels[prod]),
    n = nrow(dat),
    ordinal_omega = unname(
      semTools::compRelSEM(fit, ord.scale = TRUE)[1]
    )
  )
})

product_reliability_table <- dplyr::bind_rows(product_reliability_results) %>%
  dplyr::mutate(ordinal_omega = round(ordinal_omega, 3))

product_reliability_table
range(product_reliability_table$ordinal_omega)
```

```{r Latent product correlation reporting}
# Report the correlations among product factors from the correlated nine-factor CFA.
# These are latent correlations, not correlations among observed raw scores.
latent_product_correlation_matrix <- lavaan::lavInspect(
  fit_product_factors,
  "cor.lv"
)

latent_product_correlation_table <- as.data.frame(
  as.table(latent_product_correlation_matrix)
) %>%
  dplyr::rename(factor_1 = Var1, factor_2 = Var2, correlation = Freq) %>%
  dplyr::filter(as.numeric(factor_1) > as.numeric(factor_2)) %>%
  dplyr::arrange(dplyr::desc(correlation))

latent_product_correlation_table
range(latent_product_correlation_table$correlation)
```

```{r Higher-order scoring sensitivity}
# Extract standardized higher-order loadings.
higher_order_loading_table <- lavaan::standardizedSolution(
  fit_general_factor
) %>%
  dplyr::filter(lhs == "general_sus", op == "=~") %>%
  dplyr::transmute(
    product_factor = rhs,
    product = stringr::str_remove(rhs, "_sus$"),
    product_label = unname(product_labels[product]),
    standardized_loading = est.std
  )

higher_order_loading_table
range(higher_order_loading_table$standardized_loading)

# Compare equal and loading-weighted composites only among respondents with all
# nine product scores. This avoids changing the meaning of the composite across
# respondents, but it is a selected subgroup created by PATH product routing.
complete_nine_product_scores <- product_score_data %>%
  dplyr::filter(n_product_scores == length(product_score_names))

complete_score_matrix <- complete_nine_product_scores %>%
  dplyr::select(dplyr::all_of(product_score_names))

standardized_score_matrix <- scale(complete_score_matrix)

weight_lookup <- stats::setNames(
  higher_order_loading_table$standardized_loading,
  paste0(higher_order_loading_table$product, "_score")
)

higher_order_weights <- weight_lookup[product_score_names]

scoring_comparison <- complete_nine_product_scores %>%
  dplyr::transmute(
    equal_weight_raw = rowMeans(complete_score_matrix),
    equal_weight_standardized = rowMeans(standardized_score_matrix),
    loading_weighted_standardized = as.numeric(
      standardized_score_matrix %*% higher_order_weights /
        sum(higher_order_weights)
    )
  )

scoring_agreement_matrix <- stats::cor(scoring_comparison)
scoring_agreement_matrix

scoring_sensitivity_summary <- data.frame(
  complete_nine_n = nrow(scoring_comparison),
  raw_vs_equal_standardized = scoring_agreement_matrix[
    "equal_weight_raw", "equal_weight_standardized"
  ],
  equal_vs_loading_weighted = scoring_agreement_matrix[
    "equal_weight_standardized", "loading_weighted_standardized"
  ],
  raw_vs_loading_weighted = scoring_agreement_matrix[
    "equal_weight_raw", "loading_weighted_standardized"
  ]
) %>%
  dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))

scoring_sensitivity_summary
```

```{r General composite internal consistency}
# Evaluate internal consistency of the nine scored product subscales in the
# complete-nine subgroup. This supports only a sensitivity analysis for a simple
# composite; it does not establish routing rules or predictive validity.
general_score_model <- paste0(
  "general_score =~ ",
  paste(product_score_names, collapse = " + ")
)

fit_general_observed_score <- lavaan::cfa(
  model = general_score_model,
  data = complete_nine_product_scores,
  estimator = "MLR"
)

general_composite_reliability <- data.frame(
  n = nrow(complete_nine_product_scores),
  omega = unname(
    semTools::compRelSEM(
      fit_general_observed_score,
      ord.scale = FALSE,
      return.df = FALSE
    )[1]
  ),
  standardized_alpha = psych::alpha(
    complete_score_matrix,
    check.keys = FALSE,
    warnings = FALSE
  )$total$std.alpha
) %>%
  dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))

general_composite_fit <- lavaan::fitMeasures(
  fit_general_observed_score,
  c("cfi.robust", "tli.robust", "rmsea.robust", "srmr")
)

general_composite_fit_table <- data.frame(
  measure = names(general_composite_fit),
  value = as.numeric(general_composite_fit)
) %>%
  dplyr::mutate(value = round(value, 3))

general_composite_reliability
general_composite_fit_table
```

```{r Product-specific information functions}
# Derive all full- and brief-form curves from the same product-specific GRM.
# This puts every comparison within a product on the same theta metric.
theta_values <- seq(-3, 4, by = 0.05)
theta_matrix <- matrix(theta_values)

product_information_results <- lapply(fit_product_grm, function(x) {
  item_names <- rownames(
    mirt::coef(x$fit, IRTpars = TRUE, simplify = TRUE)$items
  )

  item_information <- vapply(
    seq_along(item_names),
    function(i) {
      mirt::iteminfo(
        mirt::extract.item(x$fit, i),
        Theta = theta_matrix
      )
    },
    numeric(length(theta_values))
  )

  colnames(item_information) <- item_names

  nextyear_position <- which(stringr::str_detect(item_names, "_nextyear$"))
  trysoon_position <- which(stringr::str_detect(item_names, "_soon$"))

  full_information <- rowSums(item_information)

  information_wide <- data.frame(
    product = x$product,
    product_label = x$product_label,
    theta = theta_values,
    full_information = full_information,
    full_standard_error = 1 / sqrt(full_information),
    nextyear_trysoon_information = rowSums(
      item_information[, c(nextyear_position, trysoon_position), drop = FALSE]
    ),
    nextyear_information = item_information[, nextyear_position],
    trysoon_information = item_information[, trysoon_position]
  )

  information_long <- information_wide %>%
    dplyr::select(
      product,
      product_label,
      theta,
      full_information,
      nextyear_trysoon_information,
      nextyear_information,
      trysoon_information
    ) %>%
    tidyr::pivot_longer(
      cols = dplyr::ends_with("information"),
      names_to = "form",
      values_to = "information"
    ) %>%
    dplyr::mutate(
      form = dplyr::recode(
        form,
        full_information = "Full four-item scale",
        nextyear_trysoon_information = "Next year + try soon",
        nextyear_information = "Next year only",
        trysoon_information = "Try soon only"
      )
    )

  list(wide = information_wide, long = information_long)
})

product_information_wide <- dplyr::bind_rows(
  lapply(product_information_results, `[[`, "wide")
)

product_information_long <- dplyr::bind_rows(
  lapply(product_information_results, `[[`, "long")
) %>%
  dplyr::mutate(
    form = factor(
      form,
      levels = c(
        "Full four-item scale",
        "Next year + try soon",
        "Next year only",
        "Try soon only"
      )
    )
  )

# Summarize where each full product scale provides maximum information.
product_information_peak_table <- product_information_wide %>%
  dplyr::group_by(product, product_label) %>%
  dplyr::slice_max(full_information, n = 1, with_ties = FALSE) %>%
  dplyr::transmute(
    product,
    product_label,
    maximum_information = full_information,
    theta_at_maximum_information = theta,
    minimum_standard_error = full_standard_error
  ) %>%
  dplyr::mutate(
    dplyr::across(
      c(maximum_information, theta_at_maximum_information, minimum_standard_error),
      ~ round(.x, 2)
    )
  )

product_information_peak_table

# Percentage of full-scale information retained from theta = 0 to 3.
# The same equally spaced grid is used for all forms, so the ratio of mean
# information is equivalent to a ratio of areas under the curves.
brief_form_information_retention <- product_information_long %>%
  dplyr::filter(theta >= 0, theta <= 3) %>%
  dplyr::group_by(product, product_label, form) %>%
  dplyr::summarise(
    mean_information_theta_0_to_3 = mean(information),
    .groups = "drop"
  ) %>%
  dplyr::group_by(product, product_label) %>%
  dplyr::mutate(
    percent_full_information_retained =
      100 * mean_information_theta_0_to_3 /
      mean_information_theta_0_to_3[form == "Full four-item scale"]
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    mean_information_theta_0_to_3 = round(mean_information_theta_0_to_3, 2),
    percent_full_information_retained = round(percent_full_information_retained, 1)
  )

brief_form_information_retention
```

```{r Figure 1 product information and standard error, fig.width=11, fig.height=9}
# Figure 1: full four-item information and conditional SE for each product.
# A separate dual-axis transformation is calculated within every product panel.
make_information_se_panel <- function(product_data) {
  # SE becomes extremely large where information approaches zero. Cap the
  # displayed SE at 2.0 so the informative region remains visible; retain the
  # uncapped values in product_information_wide for all calculations.
  product_data <- product_data %>%
    dplyr::mutate(full_standard_error_display = pmin(full_standard_error, 2))

  scaling_constant <- max(product_data$full_information, na.rm = TRUE) / 2

  ggplot2::ggplot(product_data, ggplot2::aes(x = theta)) +
    ggplot2::geom_line(
      ggplot2::aes(y = full_information, linetype = "Test information"),
      linewidth = 0.9,
      color = "black"
    ) +
    ggplot2::geom_line(
      ggplot2::aes(
        y = full_standard_error_display * scaling_constant,
        linetype = "Standard error"
      ),
      linewidth = 0.9,
      color = "gray45"
    ) +
    ggplot2::scale_y_continuous(
      name = "Information",
      sec.axis = ggplot2::sec_axis(
        ~ . / scaling_constant,
        name = "Standard error"
      )
    ) +
    ggplot2::scale_linetype_manual(
      values = c("Test information" = "solid", "Standard error" = "dashed"),
      name = NULL
    ) +
    ggplot2::coord_cartesian(xlim = c(-1, 4)) +
    ggplot2::theme_bw(base_size = 10) +
    ggplot2::theme(
      legend.position = "bottom",
      panel.grid.minor = ggplot2::element_blank(),
      plot.title = ggplot2::element_text(size = 10, hjust = 0.5)
    ) +
    ggplot2::labs(
      title = unique(product_data$product_label),
      x = expression(theta)
    )
}

information_se_panels <- split(
  product_information_wide,
  factor(
    product_information_wide$product,
    levels = names(product_items_main)
  )
) %>%
  lapply(make_information_se_panel)

figure_1_product_information <- patchwork::wrap_plots(
  information_se_panels,
  ncol = 3,
  guides = "collect"
) +
  patchwork::plot_annotation(
    title = "Test Information and Standard Error for Product-Specific Susceptibility Scales",
    subtitle = "Higher information and lower standard error indicate greater measurement precision; displayed SE is capped at 2.0"
  )

figure_1_product_information

if (!dir.exists("analysis_outputs")) dir.create("analysis_outputs")

ggplot2::ggsave(
  "analysis_outputs/Figure_1_product_information_and_SE.png",
  figure_1_product_information,
  width = 11,
  height = 9,
  dpi = 600
)

ggplot2::ggsave(
  "analysis_outputs/Figure_1_product_information_and_SE.pdf",
  figure_1_product_information,
  width = 11,
  height = 9
)
```

```{r Figure 2 within-product brief-form information, fig.width=11, fig.height=8}
# Figure 2: full and brief alternatives, calibrated within the same product GRM.
figure_2_within_product_information <- ggplot2::ggplot(
  product_information_long %>%
    dplyr::mutate(
      product_label = factor(
        product_label,
        levels = unname(product_labels[names(product_items_main)])
      )
    ),
  ggplot2::aes(
    x = theta,
    y = information,
    color = form,
    linetype = form
  )
) +
  ggplot2::geom_line(linewidth = 0.85) +
  ggplot2::facet_wrap(
    ~ product_label,
    ncol = 3,
    scales = "free_y"
  ) +
  ggplot2::coord_cartesian(xlim = c(-1, 4)) +
  ggplot2::scale_color_manual(
    values = c("black", "#0072B2", "#D55E00", "#009E73"),
    name = NULL
  ) +
  ggplot2::scale_linetype_manual(
    values = c("solid", "longdash", "dotted", "dotdash"),
    name = NULL
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(
    legend.position = "bottom",
    panel.grid.minor = ggplot2::element_blank()
  ) +
  ggplot2::labs(
    title = "Information Retained by Brief Susceptibility Forms Within Product",
    subtitle = "All alternatives within a panel use parameters from the same product-specific GRM",
    x = expression(theta),
    y = "Test information"
  )

figure_2_within_product_information

ggplot2::ggsave(
  "analysis_outputs/Figure_2_within_product_information.png",
  figure_2_within_product_information,
  width = 11,
  height = 8,
  dpi = 600
)

ggplot2::ggsave(
  "analysis_outputs/Figure_2_within_product_information.pdf",
  figure_2_within_product_information,
  width = 11,
  height = 8
)
```

```{r Export practical scoring and information results}
# Export review tables so manuscript values do not need to be transcribed by hand.
if (!dir.exists("analysis_outputs")) dir.create("analysis_outputs")

readr::write_csv(
  product_reliability_table,
  "analysis_outputs/product_subscale_reliability.csv"
)

readr::write_csv(
  latent_product_correlation_table,
  "analysis_outputs/latent_product_correlations.csv"
)

readr::write_csv(
  higher_order_loading_table,
  "analysis_outputs/higher_order_loadings.csv"
)

readr::write_csv(
  product_score_availability,
  "analysis_outputs/product_score_availability.csv"
)

readr::write_csv(
  scoring_sensitivity_summary,
  "analysis_outputs/scoring_sensitivity_summary.csv"
)

readr::write_csv(
  general_composite_reliability,
  "analysis_outputs/general_composite_reliability.csv"
)

readr::write_csv(
  general_composite_fit_table,
  "analysis_outputs/general_composite_fit.csv"
)

readr::write_csv(
  product_information_peak_table,
  "analysis_outputs/product_information_peaks.csv"
)

readr::write_csv(
  brief_form_information_retention,
  "analysis_outputs/brief_form_information_retention.csv"
)
```

# Harmonized never-user primary analyses

The primary analyses below restrict the sample to respondents with a routed
curiosity response for every included product. Under the Wave 8 questionnaire
routing, these respondents were never users of all nine included products;
nicotine-pouch eligibility additionally required prior awareness. The original
four response categories are retained for descriptive statistics. Because the
upper categories were sparse in this harmonized sample, categories 3 and 4 are
combined consistently for the primary ordinal CFA and graded response models.

```{r Harmonized never-nine primary analysis, cache=TRUE}
# The sourced analysis exports the sample-flow audit, manuscript tables,
# model diagnostics, scoring sensitivity results, and publication figures.
source("harmonized_primary_analysis.R")
```
